What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-12 android-stacks.tex
changes on abstract. introducing bug hazard

\title{What can Issues on GitHub and Google Code tell us about Exception Fragilities in Android and Java? }

@@

This paper reports a study which mined the Java stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). 
The goal of this study is to investigate  whether the exception handling logic on these stack traces
 can reveal common characteristics on the exception-related code that
are particularly fault prone and may lead to a decrease in application robustness.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Key problems identified in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues. However, inspecting the source code related to 6005 stack traces can easily become infeasible.
Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
 pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.

To perform this empirical investigation, we selected a subset of all reported root exceptions. 
This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
 Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
 It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
After doing such analysis, we could calculate the prevalence of such exceptions on issues. In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to ``programming mistakes".

Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
whose full name does not carry the reason for the exception to be thrown.
To ensure the quality of the analysis, three independent developers classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the resulting interrater agreement was high (96\%).

The results show that the programming mistakes are associated to the most reported root causes. 
Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
besides presenting an error message to the user and restarting the application~\footnote{Only 
high fault tolerant systems need to provide solutions to handle them.}.  
This finding calls the attention to the following: the high number of programming mistakes reported 
on issues are an indication that developers may be underestimating the impact that such 
mistakes have on application robustness.

\begin{table}
  \centering
  \begin{tabular}{lr}
    \hline
    \bfseries{Cause} &  \bfseries{\% Occurrences} \\
    \hline
      Programming Mistakes &  51.96\%\\ 
      %Memory Restriction   &   4.10\% \\ 
      General Exceptions   &  6.70\%\\
    \hline
 \end{tabular}
  \caption{Identifying the concerns related to root exceptions}
  \label{tab:causes}
\end{table}

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
}
\end{minipage}}

>>>>

\title{What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java? }

@@

This paper reports a study which mined the Java stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). The goal of this study is to investigate whether the analysis of stack traces information 
could reveal bug hazards related to the exception handling code that may lead to a 
decrease in application robustness. A bug hazard is a circumstance that increases the 
chance of a bug to be present in the software. Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some of the outcomes of this study includes the identification of the following 
bug hazards: (i) unexpected exception wrappings (such as Errors being wrapped in checked exceptions) -
indicating that Java's hybrid exception model become fault-prone when it is not fully used in 
accordance with its purpose; (ii) undocumented runtime exceptions thrown by both the Android 
platform and third party libraries; and (iii) undocumented checked exceptions thrown by the Android Platform.
This study provides insights on such bug hazards and the robustness threats  
they impose not only to Android apps but also to other systems based on Java exception model.

@@

%Commented 2015
%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues. However, inspecting the source code related to %6005 stack traces can easily become infeasible.
%Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
%documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
%pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.

%To perform this empirical investigation, we selected a subset of all reported root exceptions. 
%This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
%This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
%all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
%Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
% It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
%After doing such analysis, we could calculate the prevalence of such exceptions on issues. In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to %``programming mistakes".

%Table ~\ref{tab:causes} summarizes the results. It also illustrates the number of general Java exceptions - 
%whose full name does not carry the reason for the exception to be thrown.
%To ensure the quality of the analysis, three independent developers classified a randomly selected
%sample of 25 exception types (from the total 100) using the same list of concerns;
%the resulting interrater agreement was high (96\%).

%The results show that the programming mistakes are associated to the most reported root causes. 
%Although there is usually no proper handling for such exceptions~\cite{wirfs2006toward}, 
%besides presenting an error message to the user and restarting the application~\footnote{Only 
%high fault tolerant systems need to provide solutions to handle them.}.  
%This finding calls the attention to the following: the high number of programming mistakes reported 
%on issues are an indication that developers may be underestimating the impact that such 
%mistakes have on application robustness.

%\begin{table}
 % \centering
 % \begin{tabular}{lr}
 %   \hline
 %   \bfseries{Cause} &  \bfseries{\% Occurrences} \\
 %   \hline
 %     Programming Mistakes &  51.96\%\\ 
 %     %Memory Restriction   &   4.10\% \\ 
 %     General Exceptions   &  6.70\%\\
 %   \hline
% \end{tabular}
%  \caption{Identifying the concerns related to root exceptions}
%  \label{tab:causes}
% \end{table}

%\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
%The study revealed: approximately 52\% of the reported stack traces were associated to programming mistakes;
%and 27.71\% of all stack traces reported a java.lang.NullPointerException as the root cause.
%}
%\end{minipage}}

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
This identification performed in this study was based on meaning that the name of the exception holds,
its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility
issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources
of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility
~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e., java.lang.Exception,
java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).

To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
Javadoc related to each exception, and Java specification, we identified the concern
releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks.
Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.

\begin{table}
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
\hline
Programming logic (java.lang and util) & 52,0\%\\
Resources (IO, Memory, Batery) & 23,9\% \\
Security & 4,1\%\\
Concurrency & 2,9\% \\
Backward compatibility & 5,5\% \\
Specific Exceptions & 4,9\%\\
General (Error, Exception, Runtime) & 6,7\%\\
\end{tabular}
\caption{Identifying the concerns related to root exceptions}
\label{tab:tophundrend}
\end{table}
To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive
enough for covering the range of exception types analyzed.
These study showed that the programming mistakes associated to the most reported root causes.
However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
lead to a NullPointerException.
\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of exception stack traces.
51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}
